Optimal tensor denoising and completion based on ordinal observations

### 2020-01-22 ICML/miaoyan/ordinalT.tex
update

\twocolumn[
\icmltitle{Tensor denoising and completion based on ordinal observations}

@@

\begin{abstract}
Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, social network analysis, and psychological studies. We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations. Two related problems are studied, one on tensor denoising and another on tensor completion. 
%{\color{red}In contrast to the continuous-valued problems, problems with ordinal tensors exhibit an intrinsic phase transition depending on the signal-to-noise ratio. (Not sure whether this should be mentioned)}
We propose a multi-linear cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees. 
Our error bound enjoys a faster convergence rate than existing results, and we show that the proposed estimator is minimax optimal under the considered model. 
An efficient completion method is further developed, which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional tensor using only $O(Kd)$ noisy, quantized observations. We demonstrate the outperformance of our approach over previous methods through applications to the tasks of clustering and collaborative filtering.

@@

\begin{table*}[http]
\begin{center}
\includegraphics[width=17cm]{compare.pdf}
\end{center}
\vspace{-.5cm}
\caption{Comparison of various low-rank estimation methods based on categorical observations. For an order-$K$ dimensional-$(d,\ldots,d)$ ordinal tensor with $L$-level observations, we report the error bound in the recovered signal (for tensor denoising) and required sample complexity (for tensor completion) as functions of tensor dimension. {\color{red}(font in the table seems different)}}~\label{tab:compare}
\end{table*}

@@

\begin{thm}[Minimax lower bound]\label{thm:minimax} 
Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$. Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \tA\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$. Then, under the model~\eqref{eq:model},
\begin{align}\label{eq:lower}
\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
& \geq {1\over 256} \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
\end{align}
where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of the tensor dimension and rank. 
\end{thm}

@@

\begin{thm}~\label{thm:completion}
Assume the same set-up as in Theorem~\ref{thm:rate}. Suppose that we observe a subset of tensor entries $\{y_\omega\}_{\omega\in\Omega}$, where $\Omega$ is chosen at random with replacement according to a probability distribution $\Pi$. Let $\hat \Theta$ be the solution to~\eqref{eq:estimator}, and assume $r_{\max}=O(1)$. Then, with very high probability over the sampled set $\Omega$, 
\[
\PiFnormSize{}{\Theta-\hat \Theta}^2 \to 0,\quad \text{ as }\quad {|\Omega|\over \sum_k d_k} \ \to \infty. 
\]
\end{thm}
Theorem~\ref{thm:completion} shows that our estimator achieves consistency when sample size is greater than $O(\sum_k d_k)$. {\color{red} add remarks comparing with prior work, and with the degree of freedom.}

>>>>

\twocolumn[
\icmltitle{Optimal tensor denoising and completion based on ordinal observations}

@@

\begin{abstract}
Higher-order tensors arise frequently in applications such as neuroimaging, recommendation system, social network analysis, and psychological studies. We consider the problem of low-rank tensor estimation from possibly incomplete, ordinal-valued observations. Two related problems are studied, one on tensor denoising and another on tensor completion. 
%{\color{red}In contrast to the continuous-valued problems, problems with ordinal tensors exhibit an intrinsic phase transition depending on the signal-to-noise ratio. (Not sure whether this should be mentioned)}
We propose a multi-linear cumulative link model, develop a rank-constrained M-estimator, and obtain theoretical accuracy guarantees. 
Our error bound enjoys a faster convergence rate than existing results, and we show that the proposed estimator is minimax optimal under the considered model. 
An efficient completion method is further developed, which guarantees consistent recovery of an order-$K$ $(d,\ldots,d)$-dimensional tensor using only $O(Kd)$ noisy, quantized observations. We demonstrate the outperformance of our approach over previous methods on the tasks of clustering and collaborative filtering.

@@

\begin{table*}[http]
\centerline{\includegraphics[width=16cm]{compare.pdf}}
\vspace{-.4cm}
\caption{Comparison of various low-rank estimation methods based on categorical observations. For an order-$K$ dimensional-$(d,\ldots,d)$ ordinal tensor with $L$-level observations, we report the error bound in the recovered signal (for tensor denoising) and required sample complexity (for tensor completion) as functions of tensor dimension. {\color{red}(font in the table seems odd...)}}~\label{tab:compare}
\end{table*}

@@

\begin{thm}[Minimax lower bound]\label{thm:minimax} 
Assume the same set-up as in Theorem~\ref{thm:rate}, and $d_{\max}=\max_k d_k \geq 8$. Let $\inf_{\hat \Theta}$ denote the infimum over all estimators $\hat \Theta\in\tP$ based on the ordinal tensor observation $\tY\in[L]^{d_1\times \cdots \times d_K}$. Then, under the model~\eqref{eq:model},
\begin{align}\label{eq:lower}
\inf_{\hat \Theta }\sup_{\trueT\in\tP} \mathbb{P}\Big\{& \text{MSE}(\hat \Theta, \trueT) \\
& \geq {1\over 256} \min\left( \alpha^2, \ { Cr_{\max}d_{\max} \over \prod_k d_k} \right) \Big\} \geq {1\over 8},
\end{align}
where $C=C(\alpha, L, f,\mb)>0$ is a constant independent of the tensor dimension and rank. 
\end{thm}

@@

\begin{thm}~\label{thm:completion}
Assume the same set-up as in Theorem~\ref{thm:rate}. Suppose that we observe a subset of tensor entries $\{y_\omega\}_{\omega\in\Omega}$, where $\Omega$ is chosen at random with replacement according to a probability distribution $\Pi$. Let $\hat \Theta$ be the solution to~\eqref{eq:estimator}, and assume $r_{\max}=O(1)$. Then, with very high probability over the sampled set $\Omega$ ({\color{red} not over $\tY$?}), 
\[
\PiFnormSize{}{\Theta-\hat \Theta}^2 \to 0,\quad \text{ as }\quad {|\Omega|\over \sum_k d_k} \ \to \infty. 
\]
\end{thm}
Theorem~\ref{thm:completion} shows that our estimator achieves consistency when sample size is greater than $O(\sum_k d_k)$. {\color{red} add remarks comparing with prior work, and with the degree of freedom.}
